HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET BRAIN ALGORITHM

### 2012-10-31 paper.Rnw
Piles of edits incorporating more of Mallars comments and clarifications to the results and discussion section.

\newcommand{\todo}[1]{
    \textcolor{red}{#1}
}

\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET
BRAIN ALGORITHM} 
\name{  Jon Pipitone$^{\star}$ Jason P. Lerch$^{\dagger}$ Miriam Friedel$^{\dagger}$ Aristotle Voineskos$^{\star}$
         Mallar Chakravarty$^{\star}$ }

\address{$^{\star}$ Centre for Additions and Mental Health, Toronto ON, Canada \\
     $^{\dagger}$ Mouse Imaging Centre, ..., Toronto ON, Canada}

@@

\begin{abstract}
Neuroimaging research often relies on automated anatomical
segmentations of MR images of the brain. Current multi-atlased based
approaches provide accurate segmentations of brain images by
propagating region labels from manual delineations to unlabeled
images.  Unfortunately, these approaches often rely on a large number
of such manual segmentations which take time and significant expertise
to produce, neither of which may be readily available.  We present
an algorithm for the automatic segmentation of the hippocampus that
minimizes the number of atlases needed whilst still achieving nearly
the same accuracy as other multi-atlas approaches.  We perform
repeated random sub-sampling validation on the IBSR dataset to
validate our approach and find an average best case percent difference
to the multi-atlas approach of 1.9\%. \todo{i still think that his sentence
needs to be clearer; break this out into two sentences - or more- where
first you talk about the validation approach, including the different
methodologies.  in a second sentence you should discuss the results.  also
i think we need some emphasis here on how much better this approach is than
the regular naive/multi-atlas-based procedures.  there is a clear benefit to
using maget brain when you don't want to invest the time into creating a
template library.}
% fix last line: clarify "best case percent difference"

@@

In this paper we propose an automated segmentation technique to address the
issues found in existing multi-atlas-based methods  \todo{AND CREATE AND
IMPROVEMENT OVER REGULAR ATLAS-BASED PROCEDURES??}.. Principly, our method
aims to dramatically reduce the number of manually labelled atlases
necessary.  It does this by using a small atlas set to {\it generate} a
much larger ``template library'', which is then used to segment the
subjects in the same fashion as other multi-atlas-based methods.  This
approach has the additional advantage of using the unique subject
population on hand to bootstrap the segmentation analysis.

The essential insight of generating a template library is not new.
Heckemann \cite{Heckemann2006} compared generating a template library from
a single atlas to standard multi-atlas segmentation and found poor
performance and so deemed the approach as inviable.  The LEAP algorithm
\cite{Wolz2010} proceeds by iteratively segmenting the unlabelled image
most similar to the atlas library images and then incorporating the
now-labelled image into the atlas library, but requires 30 starting
atlases.  Our contribution is a method that outputs comparable segmentation
accuracy to these and other multi-atlas-based methods whilst using
significantly fewer manually created atlases.

In previous work from our group \cite{Chakravarty2011}, we explored
this same approach to the segmentation of certain subcortical structures (striatum,
globus pallidus, and thalamus) using a single histologically-derived atlas.
\todo{challenge the "MODEST INCREMENTAL ADVANCE" argument} In this work we
extend our approach to the hippocampus, and validate it by varying the
atlas and template library sizes.  Due to the small number of atlases
required, our method could easily accommodate different hippocampal
definitions.  Our aim is not to improve on segmentation accuracy beyond
existing methods, but instead to provide a method that trades off manual
segmentation expertise for computational processing time whilst providing
sufficient accuracy for clinical and research applications.

@@

We use the {\it cross-correlation weighted voting} method to both improve the
quality of the label fusion step and reduce the number of registrations
required \todo{ref}.  In this method, each template library image is ranked 
in similarity to each unlabelled image, and only the top $n$ template library
images labels are used for the voxel-wise majority vote fusion step.
Similarity scores are computed by the normalized cross-correlation of image
intensities between an unlabeled image and each template library image, after
the unlabeled image has undergone linearly registration to the template image.
The cross-correlation score is calculated over a region of interest (ROI)
on each template, representing a generous approximation of hippocampal
structure and neighbourhood, defined as the extent of all atlas labels
after linear registration to the template, dilated by three voxels.
\todo{ref Heckmann}.

\subsection{Image Processing and Registration Method}
The N3 algorithm \cite{Sled1998} is first used to minimize the intensity
nonuniformity in each of the atlases and unlabeled subject images.  Image
registration is carried out in two phases.  In the first, a 12-parameter linear
transformation (3 translations, rotations, scales, shears) is estimated between
images using an algorithm that maximizes the correlation between blurred MR
intensities and gradient magnitude over the whole brain \cite{Collins}.  In the
second phase, nonlinear registration is completed using the ANIMAL algorithm
\cite{Collins1995}: an iterative procedure that estimates a 3D deformation
field between two MR images. At first, large deformations are estimated using
blurred version of the input data. These larger deformations are then input to
subsequent steps where the fit is refined by estimating smaller deformations on
data blurred with a Gaussian kernel with a smaller FWHM. The final
transformation is a set of local translations defined on a bed of equally
spaced nodes that were estimated through the optimization of the correlation
coefficient. For the purposes of this work we used the regularization
parameters optimized in Robbins et al.\cite{Robbins2004}. It should be
noted that the MAGeT brain algorithm is not dependant on this, or any,
particular choice of registration method\cite{Chakravarty2011}.

@@

@@

We explored how varying the size of the atlas library and the template library
effects labeling accuracy.  For each parameter setting we conducted 30 rounds
of random subsampling cross-validation using the 18 manual segmented templates
from the IBSR dataset as input. In each round, atlases were randomly chosen
from the IBSR dataset and the remaining images are used both as template
library images and unlabeled subjects to be labeled using the MAGeT Brain
algorithm.  We varied the size of the atlas library from 3 to 8, and used
cross-correlation weighted label fusion to select the top $n$ candidate
templates from the remaining images.  The size of the template library ($n$)
was varied in the range $[3, 18-a]$ (where $a$ is the size of the atlas
library).

@@

\subsubsection{Comparison Approaches}
The resulting segmentations from each of our experiments is compared to two
alternative segmentation approaches. The {\it naive} approach is that in  which
a single atlas is used to segment an unlabeled subject by directly propagating
labels.  The {\it multi-atlas} approach, as described above, in which we used
cross-correlation to select the top most similar atlases to each subject a
atlas
labels are propagated to an unlabeled subject and the labels are fused using
the cross-correlation voting described above.  In total we evaluated
approximately $52,000$ segmentations for the work presented in this manuscript.

@@

Sample segmentations from one of the IBSR subjects is shown in Fig.
\ref{montage}.  The segmentations shown are varied across the template
selection for a case where 3 atlases were used.  The figure demonstrates a
reduction in false negatives (voxels labeled by the `gold-standard' only)
in the hippocampal tail, a constant \todo{HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
labeled by the MAGeT Brain method only) in the hippocampal head, and an
increase in the segmentation accuracy as the number of templates used from
the template library are increased.

\begin{figure}[h]
\begin{minipage}[b]{1.0\linewidth}
  \centering
  \includegraphics[width=\textwidth]{montage.png}
\end{minipage}
\caption{{\em Sample MAGeT Brain segmentations.} Segmentations of a single
subject are shown when 3 atlases are used, with cross-correlation label
fusion. Blue represents agreement between the gold-standard and MAGeT
Brain.  Red indicates voxels labeled only by the gold-standard, and green
indicates voxels labeled only by MAGeT Brain. Note improved accuracy as the
number of templates is increased.} 
\label{montage}
\end{figure}

In Fig. \ref{results} we compare the performance of the MAGeT Brain method
across all validations (i.e. a combination of all atlases and template
selections) against the multi-atlas method and the naive approach.
Regardless of the number of atlases and templates used in a particular
segmentation we see a marked increase in segmentation accuracy over the
average naive segmentation (the average Kappa of all segmentations derived
from all pairwise registrations in the IBSR dataset).  Clearly, including
more input atlases increases segmentation accuracy.  Surprisingly,
increasing the number of labels in the label fusion step through
cross-correlation based voting, does little to improve segmentation
accuracy (except for in the cases where 3 and 4 atlases are used).
Finally, the best-case MAGeT Brain performance scenario is when 8 atlases
are used.  While this does not reach the accuracy of the multi-atlas
segmentation, there is only a 1.9\% difference in the average Kappa values
($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
selecting the top 14 templates).  \todo{SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
library(ggplot2)
results = read.csv(file='data/results.csv')
xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
  geom_smooth(line=1.5, formula=y~poly(x,2), method='lm') + 
  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
  theme(legend.position="none") + 
  ylab('Kappa') + 
  xlab('Number of Atlases') +
  annotate("text", 16.0, 0.712, label="Naive", size=4, hjust=0)+
  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
  annotate("text", 14.6, 0.793, label="Multi-atlas", size=4, hjust=0)+
  coord_cartesian(ylim=c(0.7,0.8), xlim=c(3,18)) + 
  scale_y_continuous(breaks=seq(0,1,0.01))

@  
\end{minipage}
\caption{{\em Performance of MAGeT Brain across all validations compared to
mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
Multi-atlas and naive segmentations procedures provide upper and lower bounds
(respectively) for accuracy.  Note that improved segmentation accuracy is
achieved through increasing the number of input atlases.}
\label{results}
\end{figure}

@@

In this paper, we have demonstrated that accurate segmentations can be
achieved by simply automatically deriving a template library from a small
set of input atlases.  MAGeT Brain segmentations were compared to both
naive segmentations and multi-atlas segmentations with cross-correlation
voting.  In \cite{Lotjonen2010}, the authors report a Kappa of 0.814 on
hippocampal segmentation in the IBSR dataset using a multi-atlas approach
where atlas selection is based on the similarity between atlas and
unlabeled subject after nonlinear registration, and using STAPLE
\cite{Warfield2004} for label fusion.  We demonstrated that the trade off
in accuracy over a multi-atlas segmentation using the entire IBSR manually
segmented library is only on average 1.9\%, using only 8 randomly chosen
atlases.  Discrepancies between our results and the above results may be
due to choice of registration algorithm, regularization parameters, or
similarity metric for label fusion.  Performance may also be affected by
the variability in the IBSR data set (as previously noted by
\cite{Klein2009}).  Future work from our group will attempt to address some
of these issues.

\todo{discuss CC voting and why accuracy is monotonically increasing...
wouldn't we expect an increase and then decrease?}

>>>>

\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\magetbrain}{MAGeT brain }

\title{HIPPOCAMPAL SEGMENTATION WITH A VERY SMALL ATLAS LIBRARY: THE MAGET
BRAIN ALGORITHM} 
\name{  Jon Pipitone$^{1}$ Jason P. Lerch$^{4,5}$ Miriam Friedel$^{4,5}$
        Aristotle Voineskos$^{1,3}$ Mallar Chakravarty$^{1,2,3}$ }

\address{
$^1$ Kimel Family Translational Imaging Genetics Research Laboratory, \\
     Research Imaging Centre, Centre for Addiction and Mental Health, Toronto, ON, Canada \\
$^2$ Institute of Biomaterials and Biomedical Engineering, University of
     Toronto, Toronto, ON, Canada \\
$^3$ Department of Psychiatry, University of Toronto, Toronto, ON, Canada \\
$^4$ Program in Neuroscience and Mental Health, The Hospital for Sick Children,
     Toronto, ON, Canada \\
$^5$ Department of Medical Biophysics, University of Toronto, Toronto, ON, Canada
}

@@

\begin{abstract}
Neuroimaging research often relies on automated anatomical segmentations of MR
images of the brain. Current multi-atlas based approaches provide accurate
segmentations of brain images by propagating region labels from manual
delineations to unlabeled images. Unfortunately, these approaches often rely on
a large number of such manual-segmentated atlases which take time and
significant expertise to produce. We present an algorithm for the automatic
segmentation of the hippocampus that minimizes the number of atlases needed
whilst still achieving similar accuracy to  other multi-atlas approaches.  We
perform repeated random subsampling validation on the IBSR dataset to compare
our approach to multi-atlas segmentation using the full IBSR dataset, and to
single-atlas (model-based) segmentation. Our results show that with only 8
input atlas, \magetbrain can achieve to within 2.0\% segmentation accuracy of
the multi-atlas approach using 17 input atlases (mean $\kappa = 0.775$ vs.
$\kappa = 0.791$).

@@

In this paper we propose an automated segmentation technique to address the
above issues found in existing multi- atlas-based methods. Principly, our
method aims to dramatically reduce the number of manually labelled atlases
needed (under 10). It does this by using the small atlas library to boot-strap
a much larger "template library", which is then used to segment the subjects in
similar fashion as other multi-atlas-based methods. Our approach has the
additional advantage of using the unique subject population on hand to
initialise the segmentation analysis.

The essential insight of generating a template library is not new.  Heckemann
\cite{Heckemann2006} compared generating a template library from a single atlas
to standard multi-atlas segmentation and found poor performance and so deemed
the approach as inviable.  The LEAP algorithm \cite{Wolz2010} proceeds by
iteratively segmenting the unlabelled image most similar to the atlas library
images and then incorporating the now-labelled image into the atlas library,
but requires 30 starting atlases.  The novelty of our method is to demonstrate
the possibility of producing comparable segmentation accuracy to these and
other multi-atlas-based methods whilst using significantly fewer manually
created atlases.

In previous work from our group \cite{Chakravarty2011}, we applied \magetbrain
to the segmentation the striatum, globus pallidus, and thalamus using a single
histologically-derived atlas. In the present work we extend our approach to the
hippocampus, a much more difficult structure to segment, and perform a thorough
validation over a range of atlas and template library sizes. Due to the small
number of atlases required, our method could easily accommodate different
hippocampal definitions. Our aim is not to improve on segmentation accuracy
beyond existing methods, but instead to provide a method that trades off manual
segmentation expertise for computational processing time whilst providing
sufficient accuracy for clinical and research applications.

@@

Label fusion is performed by cross-correlation weighted voting, a strategy
weighted towards an optimal combination of subjects from the template library
which has been previously shown to improve segmentation
accuracy\cite{Aljabar2009,Collins2010}. In this method, each template library
image is ranked in similarity to each unlabelled image by the normalized
cross-correlation of image intensities after linear registration in a region of
interest (ROI) generously encompassing hippocampus.  Only the top n ranked
template library image labels are used in a voxel-wise majority vote. The ROI
is heuristically defined as the extent of all atlas labels after linear
registration to the template, dilated by three voxels\cite{Chakravarty2012}.

@@

\subsection{Image Processing and Registration Method}
The N3 algorithm \cite{Sled1998} is first used to minimize the intensity
nonuniformity in each of the atlases and unlabeled subject images.  Image
registration is carried out in two phases.  In the first, a 12-parameter linear
transformation (3 translations, rotations, scales, shears) is estimated between
images using an algorithm that maximizes the correlation between blurred MR
intensities and gradient magnitude over the whole brain \cite{Collins}.  In the
second phase, nonlinear registration is completed using the ANIMAL algorithm
\cite{Collins1995}: an iterative procedure that estimates a 3D deformation
field between two MR images. At first, large deformations are estimated using
blurred version of the input data. These larger deformations are then input to
subsequent steps where the fit is refined by estimating smaller deformations on
data blurred with a Gaussian kernel with a smaller FWHM. The final
transformation is a set of local translations defined on a bed of equally
spaced nodes that were estimated through the optimization of the correlation
coefficient. For the purposes of this work we used the regularization
parameters optimized in Robbins et al.\cite{Robbins2004}. It should be
noted that the \magetbrain algorithm is not dependant on this, or any,
particular choice of registration method\cite{Chakravarty2011}.

@@

We explored how varying the size of the atlas library and the template library
effects labeling accuracy.  For each parameter setting we conducted 30 rounds
of random subsampling cross-validation using the 18 manual segmented templates
from the IBSR dataset as input. In each round, atlases were randomly chosen
from the IBSR dataset and the remaining images are used both as template
library images and unlabeled subjects to be labeled using the \magetbrain
algorithm.  We varied the size of the atlas library from 3 to 8, and used
cross-correlation weighted label fusion to select the top $n$ candidate
templates from the remaining images.  The size of the template library ($n$)
was varied in the range $[3, 18-a]$ (where $a$ is the size of the atlas
library).

@@

\subsubsection{Comparison Approaches}
The resulting segmentations from each of our experiments is compared to two
alternative segmentation approaches. The {\it single-atlas} approach uses one
atlas to segment an unlabelled subject by directly propagating labels from
atlas by way of nonlinear registration.  For each image in the IBSR dataset we
computed the single-atlas segmentation from each of the other 17 labelled
images. For the {\it multi-atlas} approach, described above, we performed
leave-one-out validation in which each image in the IBSR dataset is segmented
using an atlas library of the remaining 17 images. To compare with the MAGeT
brain results we also varied the number of atlas images used in the label
fusion step by employing cross-correlation weighted voting.  In total we
evaluated approximately $52,000$ segmentations for the work presented in this
manuscript.

@@

Sample segmentations from a single IBSR subject compared with the gold-standard
segmentation are in Fig.  \ref{montage}.  As the size of the template library
is increased, the number of false negatives in the hippocampal tail region is
reduced.  This is correlated with an increase in segmentation accuracy.

\begin{figure}[h]
\begin{minipage}[b]{1.0\linewidth}
  \centering
  \includegraphics[width=\textwidth]{montage.png}
\end{minipage}
\caption{{\em Sample \magetbrain segmentations.} 
Segmentations of a single subject are shown when three atlases are used, with
varying template library size $n$. Blue colouring represents agreement between
the gold-standard and \magetbrain. Green colouring indicates false positive
voxels labelled by \magetbrain, and red colouring indicates false negative
voxels.
} 
\label{montage}
\end{figure}

\magetbrain shows distinct improvement in segmentation accuracy with increasing
atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
atlases (mean $\kappa = 0.775$).  This level of accuracy is within 2.0\% of the
multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
atlases, \magetbrain reaches to within 6.5\% accuracy of the multi-atlas
approach, and performs significantly better than the average single-atlas
performance. Surprisingly, our results do not show any significant improvements
in segmentation accuracy when applying cross-correlation weighted voting to
reduce the number of template labels being fused.

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
<<fig=TRUE, cached=TRUE, echo=FALSE, warnings=FALSE, >>=
library(ggplot2)
results = read.csv(file='data/results.csv')
xcorr_vote= subset(results, method=='xcorr' | method == 'multiatlas_xcorr_vote' )
qplot(x=top_n, y=k_hippocampus, data=xcorr_vote, colour=as.factor(num_atlases), geom=c('point'), alpha=I(0)) +
  geom_smooth(method="loess") + #line=1.5, formula=y~poly(x,2), method='lm') + 
  geom_hline(aes(yintercept = mean(subset(results, method=='naive')$k_hippocampus))) + 
  theme(legend.position="none") + 
  ylab('Mean similarity score (Kappa)') + 
  xlab('Number of template labels fused') +
  annotate("text", 14.3, 0.712, label="Single-atlas", size=4, hjust=0)+
  annotate("text", 15.3, 0.754, label="3 atlases", size=4, hjust=0)+
  annotate("text", 14.4, 0.766, label="4 atlases", size=4, hjust=0)+
  annotate("text", 13.5, 0.770, label="5 atlases", size=4, hjust=0)+
  annotate("text", 12.5, 0.773, label="6 atlases", size=4, hjust=0)+
  annotate("text", 11.5, 0.776, label="7 atlases", size=4, hjust=0)+
  annotate("text", 10.3, 0.779, label="8 atlases", size=4, hjust=0)+
  annotate("text", 14.5, 0.793, label="Multi-atlas", size=4, hjust=0)+
  annotate("text", 14.6, 0.787, label="17 atlases", size=4, hjust=0)+
  scale_colour_manual(values=c(
    "#D73027","#FC8D59","#FEE08B","#D9EF8B", "#91CF60","#1A9850", "#000000")) + 
  coord_cartesian(ylim=c(0.703,0.803), xlim=c(3,18)) + 
  scale_y_continuous(breaks=seq(0,1,0.01))

@  
\end{minipage}
\caption{
Mean performance of \magetbrain with varying atlas library size and number of
template labels fused. Also shown is the mean multi-atlas performance when
using an atlas library of 17 images and varying the number of labels fused, as
well as the mean performance of single-atlas segmentations.  Data is fit with
LOESS local regression smoothing lines, and include a standard deviation
ribbon.
}
\label{results}
\end{figure}

@@

In this paper, we have demonstrated that accurate segmentations can be produced
by automatically deriving a template library from a small set of input atlases.
\magetbrain segmentations were compared to both single-atlas segmentations and
multi-atlas segmentations with cross-correlation voting.  For the IBSR dataset,
on average \magetbrain achieves within 2.0\% of the segmentation accuracy of
the multi-atlas approach but requires only 8 atlases as compared to using the
entire IBSR manually segmented library of 17 atlases.

L\"{o}tj\"{o}nen et al.\cite{Lotjonen2010} report a Kappa of 0.814 on
hippocampal segmentation in the IBSR dataset using a multi-atlas approach where
atlas selection is based on the similarity between atlas and unlabeled subject
after nonlinear registration, and using STAPLE\cite{Warfield2004} for label
fusion.  Discrepancies between our results and the above results may be due to
choice of registration algorithm, regularization parameters, or similarity
metric for label fusion.  Performance may also be affected by the variability
in the IBSR data set (as previously noted by \cite{Klein2009}).  Future work
from our group will attempt to address some of these issues.

\subsection{Acknowledgements}
Computations were performed on the gpc supercomputer at the SciNet HPC
Consortium\cite{Loken2010}. SciNet is funded by: the Canada Foundation for
Innovation under the auspices of Compute Canada; the Government of Ontario;
Ontario Research Fund - Research Excellence; and the University of Toronto.

### 2012-10-31 paper.tex
Piles of edits incorporating more of Mallars comments and clarifications to the results and discussion section.

Sample segmentations from one of the IBSR subjects is shown in Fig.
\ref{montage}.  The segmentations shown are varied across the template
selection for a case where 3 atlases were used.  The figure demonstrates a
reduction in false negatives (voxels labeled by the `gold-standard' only)
in the hippocampal tail, a constant \todo{HOW DO YOU KNOW IT'S CONSTANT ??}  number of false positives (voxels
labeled by the MAGeT Brain method only) in the hippocampal head, and an
increase in the segmentation accuracy as the number of templates used from
the template library are increased.

\begin{figure}[h]
\begin{minipage}[b]{1.0\linewidth}
  \centering
  \includegraphics[width=\textwidth]{montage.png}
\end{minipage}
\caption{{\em Sample MAGeT Brain segmentations.} Segmentations of a single
subject are shown when 3 atlases are used, with cross-correlation label
fusion. Blue represents agreement between the gold-standard and MAGeT
Brain.  Red indicates voxels labeled only by the gold-standard, and green
indicates voxels labeled only by MAGeT Brain. Note improved accuracy as the
number of templates is increased.} 
\label{montage}
\end{figure}

In Fig. \ref{results} we compare the performance of the MAGeT Brain method
across all validations (i.e. a combination of all atlases and template
selections) against the multi-atlas method and the naive approach.
Regardless of the number of atlases and templates used in a particular
segmentation we see a marked increase in segmentation accuracy over the
average naive segmentation (the average Kappa of all segmentations derived
from all pairwise registrations in the IBSR dataset).  Clearly, including
more input atlases increases segmentation accuracy.  Surprisingly,
increasing the number of labels in the label fusion step through
cross-correlation based voting, does little to improve segmentation
accuracy (except for in the cases where 3 and 4 atlases are used).
Finally, the best-case MAGeT Brain performance scenario is when 8 atlases
are used.  While this does not reach the accuracy of the multi-atlas
segmentation, there is only a 1.9\% difference in the average Kappa values
($0.775$ for MAGeT Brain with 8 atlases and $0.790$ for multi-atlas after
selecting the top 14 templates).  \todo{SOMETHING IN HERE AS AN EASY WAY TO BOOST ACCURACY FROM THE "ATLAS BASED CASE WOULD BE GOOD TOO}

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
\includegraphics{paper-001}
\end{minipage}
\caption{{\em Performance of MAGeT Brain across all validations compared to
mean naive segmentation, and mean multi-atlas with cross-correlation voting.}
Multi-atlas and naive segmentations procedures provide upper and lower bounds
(respectively) for accuracy.  Note that improved segmentation accuracy is
achieved through increasing the number of input atlases.}
\label{results}
\end{figure}

>>>>

Sample segmentations from a single IBSR subject compared with the gold-standard
segmentation are in Fig.  \ref{montage}.  As the size of the template library
is increased, the number of false negatives in the hippocampal tail region is
reduced.  This is correlated with an increase in segmentation accuracy.

\begin{figure}[h]
\begin{minipage}[b]{1.0\linewidth}
  \centering
  \includegraphics[width=\textwidth]{montage.png}
\end{minipage}
\caption{{\em Sample \magetbrain segmentations.} 
Segmentations of a single subject are shown when three atlases are used, with
varying template library size $n$. Blue colouring represents agreement between
the gold-standard and \magetbrain. Green colouring indicates false positive
voxels labelled by \magetbrain, and red colouring indicates false negative
voxels.
} 
\label{montage}
\end{figure}

\magetbrain shows distinct improvement in segmentation accuracy with increasing
atlas library size (Fig. \ref{results}), reaching its highest accuracy using 8
atlases (mean $\kappa = 0.775$).  This level of accuracy is within 2.0\% of the
multi-atlas approach (mean $\kappa = 0.791$) using 17 atlases. With only 3
atlases, \magetbrain reaches to within 6.5\% accuracy of the multi-atlas
approach, and performs significantly better than the average single-atlas
performance. Surprisingly, our results do not show any significant improvements
in segmentation accuracy when applying cross-correlation weighted voting to
reduce the number of template labels being fused.

\begin{figure}
\begin{minipage}[b]{1.0\linewidth}
  \centering
  %\includegraphics[width=\textwidth]{results}
\includegraphics{paper-001}
\end{minipage}
\caption{
Mean performance of \magetbrain with varying atlas library size and number of
template labels fused. Also shown is the mean multi-atlas performance when
using an atlas library of 17 images and varying the number of labels fused, as
well as the mean performance of single-atlas segmentations.  Data is fit with
LOESS local regression smoothing lines, and include a standard deviation
ribbon.
}
\label{results}
\end{figure}
